Data Flow Analysis\\in the Presence of Correlated Calls

### 2014-12-08 content/background.tex
comments

Formally, a data-flow analysis problem is suitable for an IFDS analysis if it can be encoded as an IFDS problem
\begin{equation}
  G^*,\,D,\,F,\,M_F,\,\sqcap)\,,
\end{equation} where $G^*=(N^*,\,E^*)$ is the supergraph of the input program with nodes $N^*$ and edges $E^*$, $D$ is a finite set of \textit{data-flow facts}, $F$ is a set of distributive dataflow functions of type $2^D\to2^D$, $M_F:\,E^*\to F$ is a function that maps supergraph edges to dataflow functions, and $M_F$ is extended to paths by composition\footnote{%
  Let $A$ be a set and $f:\,E^*\to(A\to A)$ a function from supergraph edges to functions on $A$.
  We say that $f$ is extended to  paths by composition to denote that for a path $q$ consisting of the edges $e_1,\,\dots,\,e_k$, $f(q)=f(e_k)\circ\ldots\circ f(e_1)\circ\id$.
}. The \textit{meet operator} $\sqcap$ is either union or intersection.

@@

In this way, the IFDS algorithm reduces the input data-flow problem to a graph-reachability problem.
\begin{example}
  In a taint analysis, $D$ is the set of variables in the program. If a fact $d\in D$ is reachable at a given node, then the variable is considered secret at that node. Otherwise, it is considered not secret. The question ``is $d$ secret at node $n$?'' becomes ``is there a realizable path from (\textsf{start}$_\texttt{main}$,\,\textbf{0}) to $(n,\,d)$?''.
\end{example}
\begin{example} 
  In an available-expressions analysis, $D$ is the set of all expressions in the program. If an expression $d\in D$ is reachable at a certain node, it means that it does not need to be recomputed at that node. 
\end{example}
\begin{example}
  The exploded supergraph for Listing~\ref{list:ccexample} is shown in Figure~\ref{fig:cc_edgefn_example}.
  We can see that there is a realizable path
  from the start node of the exploded graph to the variable \verb's' at~the~node~$\textsf{print(s)}$ in the \verb'B.bar' method. This means that at that node, \verb's' is considered secret.

@@

To convert a representation relation $R_f$ back into the original flow function $f$, we can use the \textit{interpretation} function $\denote {R_f}$:
\begin{align}
  f&=\denote{R_f}\notag\\
  &=\lambda D_1\,.\,\left(\left\{d_2\ |\ \exists d_1\in D_1:\,(d_1,\,d_2)\in R_f\right\}\cup\{d_2\ |\ (\mathbf0,\,d_2)\in R_f\}\right)
  \setminus\{\mathbf0\}\,.
\end{align}

@@

The overall idea behind computing the lattice elements $l_{n,\,d}$ is the following. For each inter-procedurally realizable path
\[
  p=\left[(\startmain,\,\Lambda),\,(n_1,\,d_1),\,\dots,\,(n_k,\,d_k)\right]
\]
 that starts with the entrypoint of the exploded supergraph, we compute the function $f_p$ that corresponds to $p$. The micro function consists of the composition of all individual micro functions with which the edges of $p$ are labeled:
\begin{align}
  f_p=\textsf{EdgeFn}((n_{k-1},\,d_{k-1}),\,(n_k,\,d_k))\,\circ\ldots\circ
      \textsf{EdgeFn}(\startmain,\,\Lambda),\,(n_1,\,d_1))\,.
\end{align}

@@

Let $Q$ be the set of paths that start at the entry point and end at the given node $(n,\,d)$.

The lattice element $l_{n,\,d}$ is the meet of the lattice elements corresponding to all the paths in $Q$:
\[
  l_{n,\,d}=\bigsqcap_{q\in Q}l^q_{n,\,d}\,.
\]

>>>>

Formally, a data-flow analysis problem is suitable for an IFDS analysis if it can be encoded as an IFDS problem
\begin{equation}
  (G^*,\,D,\,F,\,M_F,\,\sqcap)\,,
\end{equation} where $G^*=(N^*,\,E^*)$ is the supergraph of the input program with nodes $N^*$ and edges $E^*$, $D$ is a finite set of \textit{data-flow facts}, $F$ is a set of distributive dataflow functions of type $2^D\to2^D$, $M_F:\,E^*\to F$ is a function that maps supergraph edges to dataflow functions, and $M_F$ is extended to paths by composition\footnote{%
  Let $A$ be a set and $f:\,E^*\to(A\to A)$ a function from supergraph edges to functions on $A$.
  We say that $f$ is extended to  paths by composition to denote that for a path $q$ consisting of the edges $e_1,\,\dots,\,e_k$, $f(q)=f(e_k)\circ\ldots\circ f(e_1)\circ\id$.
}. The \textit{meet operator} $\sqcap$ is either union or intersection.

@@

In this way, the IFDS algorithm reduces the input data-flow problem to a graph-reachability problem.
\begin{example}
  In a taint analysis, $D$ is the set of variables in the program. If a fact $d\in D$ is reachable at a given node, then the variable is considered secret at that node. Otherwise, it is considered not secret. The question ``is $d$ secret at node $n$?'' becomes ``is there a realizable path from (\textsf{start}$_\texttt{main}$,\,\textbf{0}) to $(n,\,d)$?''.
\end{example}
\commentout{\begin{example} 
  In an available-expressions analysis, $D$ is the set of all expressions in the program. If an expression $d\in D$ is reachable at a certain node, it means that it does not need to be recomputed at that node. 
\end{example}}
\begin{example}
  The exploded supergraph for Listing~\ref{list:ccexample} is shown in Figure~\ref{fig:cc_edgefn_example}.
  We can see that there is a realizable path
  from the start node of the exploded graph to the variable \verb's' at~the~node~$\textsf{print(s)}$ in the \verb'B.bar' method. This means that at that node, \verb's' is considered secret.

@@

\begin{mdelete}\mtodo{We never use the interpretation function}
To convert a representation relation $R_f$ back into the original flow function $f$, we can use the \textit{interpretation} function $\denote {R_f}$:
\begin{align}
  f&=\denote{R_f}\notag\\
  &=\lambda D_1\,.\,\left(\left\{d_2\ |\ \exists d_1\in D_1:\,(d_1,\,d_2)\in R_f\right\}\cup\{d_2\ |\ (\mathbf0,\,d_2)\in R_f\}\right)
  \setminus\{\mathbf0\}\,.
\end{align}
\end{mdelete}

@@

The overall idea behind computing the lattice elements $l_{n,\,d}$ is the following. For each inter-procedurally realizable path
\[
  p=\left[(\startmain,\,\Lambda),\,(n_1,\,d_1),\,\dots,\,(n_k,\,d_k)\right]
\]
 that starts with the entrypoint of the exploded supergraph, we compute the micro function $f_p$ that corresponds to $p$. The micro function consists of the composition of all individual micro functions with which the edges of $p$ are labeled:
\begin{align}
  f_p=\textsf{EdgeFn}((n_{k-1},\,d_{k-1}),\,(n_k,\,d_k))\,\circ\ldots\circ
      \textsf{EdgeFn}(\startmain,\,\Lambda),\,(n_1,\,d_1))\,.
\end{align}

@@

Let $Q$ be the set of paths that start at the entry point and end at the given node $(n,\,d)$.
The lattice element $l_{n,\,d}$ is the meet of the lattice elements corresponding to all the paths in $Q$:
\[
  l_{n,\,d}=\bigsqcap_{q\in Q}l^q_{n,\,d}\,.
\]
